A pair of co-moving, Sun-like stars with different chemical abundances:\\ a possible signature of copious rocky accretion (needs work)

### 2017-04-16 paper/ms.tex
add text to data section and new figures

\begin{abstract}
  We report and discuss the discovery of a co-moving pair of bright solar-type
  stars, HD 240429 and HD 240430, with significant differences in their chemical
  abundance patterns.
  The two stars are co-spatial ($|\Delta \bs{x}| < XX~\pc$) at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities
  ($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence), as inferred from \gaia\
  \tgas\ proper motions and high resolution radial velocity measurements.
  Both stars are young with ages $\approx$$500~{\rm Myr}$ as estimated from
  their lithium abundances, favoring the idea that they were born together and
  are or were recently a bound, wide binary system.
  However, the metallicities and chemical abundances of refractory
  elements of the two stars are very different, challenging this simple picture.
  In order to reconcile the seemingly contradictory observations,
  we consider scenarios in which the pair is unrelated despite
  its current spatial and kinematic configuration, and
  those in which the pair formed together but may now have different chemistry.
  In the latter, we consider the possibility that the more metal-rich of the two,
  HD 240430, accreted $\approx$10's of Earth masses of protoplanetary, rocky material,
  selectively enhancing the refractory elements.
  Direct-imaging and precision radial velocity follow-up to study the existence
  or properties of planetary systems around these stars would help distinguish
  these scenarios.
\end{abstract}

@@

In this section, we describe the data relevant to our discussion of the origin of the system.
The co-moving pair was initially discovered in our recent search
for co-moving pairs using the proper motions from Gaia DR1.
We refer the readers to \citealt{2016arXiv161202440O} and briefly describe the method here.
For a given pair, we performed a hypothesis test between the two hypotheses that
the two stars share the same 3D velocity vector and that they have
two independent 3D velocity vectors.
In the catalog of co-moving pairs (\citealt{2016arXiv161202440O}),
the pair that we present here was assigned a group id of 1199,
and the marginalized likelihood ratio (Bayes factor)
between the two hypotheses is $\ln{\mathcal{L}_1/\mathcal{L}_2} = 8.52$,
well above a relatively conservative cut of 6 that we adopted.
Additionally, we have checked that we do not find any possible additional co-moving companions
to this pair even when we lower the selection threshold to less conservative but still reasonable values.
The pair is in fact classified as a visual double star system independently
in Washington Double Star Catalog (\cite{2001AJ....122.3466M}).

In a separate effort to study detailed chemical abundances of
potential planet-hosting stars, \citet{2016ApJS..225...32B}
measured abundances of 15 different elements
(C, N, O, Na, Mg, Al, Si, Ca, Ti, V, Cr, Mn, Fe, Ni, Y)
of these two stars as well as high precision radial velocities
with high resolution spectroscopy.
We provide all measurements relevant to the present discussion in Table~\ref{tab:t2}.

@@

We first confirm that the two stars are indeed co-moving with very similar
velocities using the radial velocities.
\todo{APW: discuss kinematics, posterior of separation/dV}

>>>>

\begin{abstract}
  We report and discuss the discovery of a co-moving pair of bright solar-type
  stars, HD 240429 and HD 240430, with significant differences in their chemical
  abundance patterns.
  The two stars are co-spatial ($|\Delta \bs{x}| < XX~\pc$) at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities
  ($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence), as inferred from \gaia\
  \tgas\ proper motions and high resolution radial velocity measurements.
  Both stars are young with ages $\approx$$500~{\rm Myr}$ as estimated from
  their lithium abundances, favoring the idea that they were born together and
  are or were recently a bound, wide binary system.
  However, the metallicities and chemical abundances of refractory
  elements in the two stars are very different, challenging this simple picture.
  In order to reconcile the seemingly contradictory observations,
  we consider scenarios in which the pair is unrelated despite
  its current spatial and kinematic configuration, and
  those in which the pair formed together but may now have different chemistry.
  In the latter, we consider the possibility that the more metal-rich of the two,
  HD 240430, accreted $\approx$10's of Earth masses of protoplanetary, rocky material,
  selectively enhancing the refractory elements.
  Direct-imaging and precision radial velocity follow-up to study the existence
  or properties of planetary systems around these stars would help distinguish
  these scenarios.
\end{abstract}

@@

Though previously recognized as a visual double star system
(\citealt{2001AJ....122.3466M}), HD 240429 and HD 240430 were identified as a
candidate co-moving star pair in our recent search for co-moving stars using the
proper motions and parallaxes from the \tgas catalog, a component of \gaia\ \dr\
(the astrometric measurements are listed in \tablename~\ref{tab:t2}).
We refer the readers to this previous work (\citealt{2016arXiv161202440O}) for a
full explanation of the methodology behind this search and only briefly describe
the method here.
For a given pair, we compute a marginalized likelihood ratio between the
hypotheses (1) that a given pair of stars share the same 3D velocity vector, and
(2) that they have independent 3D velocity vectors, both given only observations
of two components of the velocities (proper motions).
We then select a sample of high-confidence co-moving star pairs by making a
conservative cut on this likelihood ratio.
In the resulting catalog of co-moving pairs (\citealt{2016arXiv161202440O}),
the pair presented in this paper was assigned a group id of 1199,
and the marginalized likelihood ratio (Bayes factor)
between the two hypotheses is $\ln{\mathcal{L}_1/\mathcal{L}_2} = 8.52$,
well above the adopted cut value of 6.
We have checked that we do not find any possible additional co-moving companions
by lowering the likelihood ratio cut for the stars around this pair.

In a separate effort to study detailed chemical abundances of
potential planet-hosting stars, \citet{2016ApJS..225...32B}
obtained high resolution spectroscopy for both HD 240429 and HD 240430.
\todo{from where? mention briefly Keck HIRES}
The resulting measurements from this effort include abundances measured for 15
checmical species (C, N, O, Na, Mg, Al, Si, Ca, Ti, V, Cr, Mn, Fe, Ni, Y) as
well as high precision radial velocities.
Other stellar parameters derived from these spectra are included in
\tablename~\ref{tab:t2}.

Combining these precise radial velocities with the \gaia\ astrometry, we can
compare differences between the inferred 6D phase-space properties for the
two stars.
We start by generating posterior samples over the Heliocentric distance, $r$,
tangential velocities, $(v_\alpha, v_\delta)$, and radial velocity, $\hat{v}_r$,
given the observed parallax, $\pi$, and proper motion components,
$(\mu_{\alpha^*}, \mu_\delta)$, and radial velocity, $v_r$.
By defining the vectors
\begin{eqnarray}
  \vec{y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          \pi &
          \mu_{\alpha^*} &
          \mu_\delta &
          v_r
        \end{array}
      \right)}\\
  \hat{\vec{y}} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          r^{-1} &
          r^{-1}\,v_\alpha &
          r^{-1}\,v_\delta &
          \hat{v}_r
        \end{array}
      \right)}
\end{eqnarray}
and by considering the covariance matrix $\mat{C}$ between the observed
quantities (provided by \tgas\ and extended to include the uncorrelated radial
velocity uncertainty), the likelihood can be written
\begin{eqnarray}
p(\pi, \mu_{\alpha^*}, \mu_\delta \given r, v_\alpha, v_\delta) &=&
  \left[\det\left(\frac{\mat{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\vec{y} - \hat{\vec{y}} \right)} \,
    \mat{C}^{-1} \,
    \left(\vec{y} - \hat{\vec{y}} \right) \right] \label{eq:likefn} \quad .
\end{eqnarray}
We adopt a uniform space density prior for the distance and an isotropic
Gaussian for any velocity component, $v$, with a dispersion $\sigma_v=25~\kms$
\begin{eqnarray}
p(r) &=&
  \begin{cases}
    \frac{3}{r_{\rm lim}^3} \, r^2 & \text{if } 0 < r < r_{\rm lim}\\
    0              & \text{otherwise}
  \end{cases}\\
p(v) &=& \frac{1}{\sqrt{2\pi}\,\sigma_v} \,
  \exp\left[-\frac{1}{2} \, \frac{v^2}{\sigma_v^2} \right] \quad .
\end{eqnarray}
For each of the two stars, we use \project{emcee}
(\citealt{Foreman-Mackey:2013}) to generate posterior samples in $(r, v_\alpha,
v_\delta, \hat{v}_r)$ by running 64 walkers for 4608 steps and discarding the
first 512 steps as the burn-in period.
For each sample, we convert the heliocentric phase-space coordinates into
Galactocentric coordinates assuming the Sun's position and velocity are $x_\odot
= (-8.3, 0, 0)~{\rm kpc}$ and $v_\odot = (10, 244, 7.17)~\kms$ (\citealt{bovy,
etc.}).
\figurename~\ref{fig:dxdv} shows differences in posterior samples converted to
Galactocentric phase-space coordinates for the two stars.
In all components, the pair is consistent with being co-moving and close in
separation; the sky-projected separation of the two stars is $\approx$$0.01~\pc$
but the median 3-space separation is $\approx$$0.6~\pc$.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{dx_dv_posterior.pdf}
  \end{center}
  \caption{%
    Differences in posterior samples over Galactocentric phase-space coordinates
    for the two stars HD 240429 and HD 240430.
    \label{fig:dxdv}}
\end{figure}

\figurename~\ref{fig:orbit} shows an orbit computed from the median posterior
sample over 6D phase-space coordinates for HD 240429 (black line), integrated
in a Milky Way-like gravitational potential (\cite{Gala:2017}). \todo{APW:
explain}
The Sun's orbit in this same Milky Way model is under-plotted for comparison
(grey line); the orbits of the stars in this pair have large scale-heights
relative to the Sun's orbit despite being young.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{orbits.pdf}
  \end{center}
  \caption{%
    Galactic orbits computed for HD 240429 (black) and the Sun (grey).
    For HD 240429, the initial conditions are set to the median posterior
    sample over the inferred 6D phase-space coordinates.
    The orbits are computed by integrating backwards from the present-day
    positions for $2.5~{\rm Gyr}$ with a time-step of $0.5~{\rm Myr}$ using the
    Leapfrog integration scheme implemented in \project{Gala}
    (\citealt{Gala:2017}). \todo{APW: citation...}
\label{fig:orbit}}
\end{figure}

@@
